CF Registry Analysis

### 2024-04-29 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
incidence tables updated

# Find each person's entry date
annual = annual %>% group_by(eDWID) %>%
  mutate(entry_date = min(ReviewYear))

@@

# Approximate year of birth
by <- annual %>%
  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
  count(birth_year) %>%
  slice_min(birth_year) %>%
  select(-n)
annual <- left_join(annual, by, by = join_by(eDWID))

@@

@@

annual$age_group = droplevels(annual$age_group)
annual = annual %>% 
  mutate(tstop = ReviewYear) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]

@@

  ppFEV1_cat = "FEV1 percent predicted",

@@

  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year",

@@

  age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"

@@

@@

```{r}
t2004 = annual %>% select()
```

```{r}
# Use flexsurv to fit a survival model that can be passed to rprev
llog <- flexsurv::flexsurvreg(Surv(tstop, as.numeric(CFRD)==2) ~ 1, 
                              data=prevdata, dist='llogis')
prev_llog <- prevalence(index='2004-01-01', 
                        num_years_to_estimate=1, 
                        data=annual, 
                        inc_formula = entrydate ~ 1,
                        surv_model=llog,
                        death_column = "first_cfrd",
                        incident_column = "first_cfrd",
                        N_boot = 100)
```

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ 1,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ age_group,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, as.numeric(CFRD)==2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, id = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

@@

```{r}
annual_split <- survSplit(Surv(tstart, tstop, as.numeric(CFRD)==2) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
```

>>>>

@@

@@

annual$tstop <- ymd(paste0(annual$ReviewYear, "-12-31"))
annual$tstart <- ymd(paste0(annual$ReviewYear, "-01-01"))
annual <- annual %>%
  mutate(entry_date = tstart[1]) %>%
  ungroup()

@@

annual$age_group <- droplevels(annual$age_group)

@@

  ppFEV1_cat = "FEV1 percent predicted category",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  A_FVC_pct_predicted = "FVC percent predicted",

@@

  A_mycocultureresults3 = "Microorganisms in myco culture", eDWID = "ID",

@@

  Age_YrEnd = "Age", bmi_perc = "BMI %ile"

@@

label(annual[, which(colnames(annual) %in% names(labels))]) <-
  labels[colnames(annual)[which(colnames(annual) %in% names(labels))]]

@@

## Demographics of Incident CFRD cases versus non-CFRD cases in 2004, 2011, and 2018

```{r}
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease, A_FEV1_pct_predicted,
    A_FVC_pct_predicted, NumPulmExacerbation, A_corticosteroids1
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
# Whole cohort
og <- as.numeric(ymd("2003-01-01"))
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ 1,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T, fun = "event",
  xlab = "Days from 2003-01-01"
)
cum_inc <- tidy(km_fit)
cum_inc$time <- as.Date(cum_inc$time + og)
kable(cum_inc)
# By age group
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ age_group,
  data = annual, id = eDWID
)
ggsurvplot(km_fit,
  data = annual, risk.table = T, cumevents = T, fun = "event",
  xlab = "Days from 2003-01-01"
)
km_table <- tidy(km_fit)
km_table$time <- as.Date(km_table$time + og)
km_table <- split.data.frame(km_table, km_table$strata)
```

```{r results='asis'}
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  print(kable(t, caption = name))
}))
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, id = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

@@

```{r}
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = og
  ) ~ .,
  data = annual, cut = c(2008, 2013, 2018), episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), event)"
```
